---
title: "Your Spotify Data from the Mood Music Study"
output: 
  flexdashboard::flex_dashboard:
params: 
  file_path: ""
theme: spacelab
---

<style>                     
.navbar {
  background-color:grey;
  border-color:grey;
}
.navbar a:hover, .navbar a:focus {
  background-color: darkgrey !important;
}
.navbar li.active a {
  background-color: darkgrey !important;
}
.navbar-brand {
  color:white!important;
}
.navbar-inverse .navbar-toggle:hover, .navbar-inverse .navbar-toggle:focus {
  background-color: darkgrey !important;
}
.navbar-inverse .navbar-collapse, .navbar-inverse .navbar-form {
    border-color: transparent;
}

</style>  

```{r setup, include=FALSE, echo=FALSE}
library(flexdashboard)
library(dplyr)            # Data manipulation (the %>% symbol)
library(tidyr)            # Data tidying (e.g. pivot_longer)
library(ggplot2)          # Data plotting
library(lubridate)        # Date time manipulation
library(formattable)      # Cool tables
library(plotly)           # Interactive graphs
library(RColorBrewer)     # Nice colours
library(cowplot)          # For theme_minimal_hgrid
library(forcats)          # Factor operations
library(scales)           # % in value box
library(NbClust)          # Algorithmically getting the optimum number of clusters
library(factoextra)       # Handy functions for clustering

# Read file
spotify <- read.csv(file = params$file_path, sep = '\t', fill = TRUE, header = TRUE, na.strings="NA")

# Set up date variables
spotify <- spotify %>% mutate(datetime = lubridate::ymd_hm(endTime),
                              wday = as.factor(wday(datetime, label=TRUE)),
                              hour = hour(datetime),
                              date = as_date(datetime),
                              month = as.factor(format(datetime, "%b")),
                              year = year(datetime),
                              yr_mnth = format(as.Date(datetime), "%Y-%m"))

# Relevel the months
spotify <- spotify %>%
  mutate(wday = forcats::fct_relevel(wday, "Mon", "Tue", "Wed", "Thu", "Fri", "Sat", "Sun")) %>%
  mutate(month = forcats::fct_relevel(month, "Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul",
                                       "Aug", "Sep", "Oct", "Nov", "Dec"))

# Useful list of the numeric features in the dataset
music_features <- c('danceability', 'energy', 'loudness', 'speechiness', 'acousticness', 'instrumentalness', 'liveness', 'valence')

```


Headlines {data-orientation=rows}
=====================================

Row
-------------------------------------

```{r descriptive, include=FALSE}
date1 <- min(spotify$date)
date2 <- max(spotify$date)
unique_tracks <- length(unique(spotify[["trackID"]]))
total_listens <- nrow(spotify)
listening_time <- sum(spotify$msPlayed) / (1000 * 3600) # Thousandth of a second, *3600 for hours. 

# Remove unknowns from now on as they bias the output. 
spotify <- spotify %>% filter(artistName != "Unknown Artist" | trackName != "Unknown Track")

repetition <- (1 - (unique_tracks/total_listens))
```

<!-- Between `r date1` and `r date2` you listened to  -->

### Total Listens

```{r}
valueBox(total_listens, icon='fas fa-music', color="#66C2A5")
```

### Hours of Music Listened To

```{r}
valueBox(round(listening_time), icon="fas fa-headphones", color="#FC8D62")
```

### Unique Tracks Listened To

```{r}
valueBox(round(unique_tracks), icon="fa-compact-disc", color="#8DA0CB")
```

### Of Your Songs Were Repeat Listens

```{r}
valueBox(scales::percent(round(repetition, 2)), icon="fas fa-redo-alt", color="#E78AC3")
```

<!-- Between `r date1` and `r date2` you listened to `r total_listens` tracks on Spotify, totalling `r round(listening_time)` hours! Out of all those listens, `r unique_tracks` tracks were unique, which means that `r round(repetition)`% of the things you listened to were things you had listened to before.  -->

Row
-------------------------------------

### Your top song each month {.no-title}

These were your top songs each month:

```{r tracks_by_month, echo=FALSE, warning=FALSE, message=FALSE}
tracks_by_month <- spotify %>% 
  select('yr_mnth', 'artistName','trackName') %>%  
  count(yr_mnth, artistName, trackName) %>%  # Count how many occurrences there are of each artist
  ungroup() %>%
  group_by(yr_mnth) %>% 
  top_n(1, n)

formattable(tracks_by_month,
                          align = c("l", "l", "l", "r"),
                          list(
                            n = color_bar("#FC8D62")
                            ),
                          col.names = c("Year-Month", "Artist", "Track", "Times Played")
                          )

```


Row {data-height=200}
-------------------------------------

### Your Top 5 Artists {.no-title}

Your top 5 artists overall were:
```{r artist_freqs, echo=FALSE, warning=FALSE, message=FALSE}
artist_freqs <- spotify %>% 
  select('artistName') %>%  # Select the artist name column
  count(artistName) %>%  # Count how many occurrences there are of each artist
  arrange(desc(n)) # Arrange them in descending order

# Lets print the top 10...
formattable(artist_freqs[0:5,], 
                          align = c("l", "r"),
                          list(
                            n = color_bar("#FFD92F")
                            ),
                          col.names = c("Artist", "Times Played")
                          )
```


Over Time {data-orientation=rows}
=====================================

Row
-------------------------------------

### How did your music change over time? {.no-title}

```{r song_features_data, include=FALSE}
# Select these features and drop na values
spotify_features <- spotify %>%
  select(all_of(music_features), 'date') %>%
  drop_na()

# Scale all varbs so min is 0 and max is 1
range0to1 <- function(x){(x-min(x))/(max(x)-min(x))}
spotify_features[, music_features] <- sapply(spotify_features[, music_features], range0to1)

# Transform to long for plotting
long_music_feats <- spotify_features %>%
  pivot_longer(cols=all_of(music_features), names_to="feature", values_to="value")

```

<!-- Spotify has lots of information about the features of the songs that you have listened to. Here we've plotted each of the numeric features on a graph to show how they have changed over time. You can interact with this graph by hovering over the lines or zooming in. Double-click to set it back to the default.  -->

```{r song_features_plot, echo=FALSE, warning=FALSE, message=FALSE}

#Set Up the ggplot
gg <- ggplot(data=long_music_feats,
             aes(x=date,
                 y=value,
                 color=feature,
                 group = feature)) +
  geom_smooth(se=FALSE) +
  theme_minimal_hgrid(12) +
  scale_colour_brewer(palette="Set2") +
  xlab("") +
  ylab("Feature Value") +
  labs(title = "Music features over time",
       colour = "") +
  theme(legend.position="bottom")

fig <- ggplotly(gg)

text_x <- paste("Date: ", lubridate::as_date(fig$x$data[[2]]$x))


fig %>%
    style(text = c(text_x, "feature"))
```

Row
-------------------------------------

<!-- ### When did you listen to the most music? -->

<!-- You see how many hours of music you listened to over time, for example by the month of the year, or by the day of the week.  -->

### Listening Hours By Month {.no-title}

```{r hours_by_month, echo=FALSE, warning=FALSE, message=FALSE}

hours_by_month <- spotify %>% select(msPlayed, yr_mnth) %>%
  group_by(yr_mnth) %>%
  summarise(msPlayed = sum(msPlayed)) %>%
  mutate(hoursPlayed = msPlayed / (1000 * 3600) )

hours_by_month_plt <- ggplot(hours_by_month, aes(x=yr_mnth, y=hoursPlayed)) +
  geom_bar(stat="identity", fill = "#66C2A5") +
  theme_minimal_hgrid(12) +
  xlab("") +
  ylab("Hours of Listening Time") +
  labs(title = "Total listening hours by month") +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5))

ggplotly(hours_by_month_plt)
```

### Listening Hours By Weekday {.no-title}

```{r hours_by_weekday, echo=FALSE, warning=FALSE, message=FALSE}

hours_by_wday <- spotify %>% select(msPlayed, wday) %>%
  group_by(wday) %>%
  summarise(msPlayed = sum(msPlayed)) %>%
  mutate(hoursPlayed = msPlayed / (1000 * 3600) )

hours_by_wday_plt <- ggplot(hours_by_wday, aes(x=wday, y=hoursPlayed)) +
  geom_bar(stat="identity", fill="#FFD92F") +
  xlab("") +
  ylab("Hours of Listening Time") +
  labs(title = "Total listening hours by weekday") +
  theme_minimal_hgrid(12) +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5))

ggplotly(hours_by_wday_plt)

```

Over Space {data-orientation=rows}
=====================================

Row {data-height=55}
-------------------------------------

### How do your songs group together? 

<!-- How does it work? First, the analysis groups together songs that are similar based on the features of the song like those in the previous plot, such as *acousticness* or *valence* (this uses a clustering algorithm called 'k-means'). The analysis then reduces all the information about the song that it can to two variables, so that they can be plotted on a 2-d graph (this is known as 'dimension reduction'). The result is the graph below, where each of your songs has been assigned to a group, given by the colour, and put somewhere on the graph.  -->

```{r k_means_clustering, echo=FALSE, warning=FALSE, message=FALSE}

# Create a subset with numeric features and the names we eventually want as tooltips as the index.
spotify_num <- spotify %>%
  select(trackName, artistName, all_of(music_features)) %>%
  mutate(name = sprintf("Track: %s<br>Artist: %s", trackName, artistName)) %>%
  distinct() %>%
  drop_na() %>%
  select(-c(trackName, artistName))

# Make the trackID the index
rownames(spotify_num) <- spotify_num$name
spotify_num$name <- NULL # Delete trackID as a column, it's the index now.

# Now scale
spotify_num <- spotify_num %>%
  scale()

# Get the best number of clusters algorithmically
n_clusters <- NbClust(spotify_num, min.nc = 2, method="kmeans", index="rubin")
n_clusters <- n_clusters$Best.nc[[1]]
```

Below are all your songs mapped onto a 2-D space The songs have been grouped into the best fitting `r n_clusters` groups based on their musical features, and each of these is a different colour on the graph. The closer the songs are on the graph, the more similar their features are. <br>

This plot is *interactive*. You can hover over each point to see the track and artist name. Zoom in and out using the buttons in the top right hand corner! 

What do you think? Do the groups make sense?

Row
-------------------------------------

### How do your songs group together? {.no-title}

```{r k_means_clustering_plot, echo=FALSE, warning=FALSE, message=FALSE}
# Run the kmeans algorithm
km <- kmeans(spotify_num, n_clusters, nstart = 25)

# Dimension reduction using PCA
res.pca <- prcomp(spotify_num,  scale = TRUE)
# Coordinates of individuals
ind.coord <- as.data.frame(get_pca_ind(res.pca)$coord)
# Add clusters obtained using the K-means algorithm
ind.coord$cluster <- factor(km$cluster)

cluster_plot <- ggplot(ind.coord,
                       aes(x = Dim.1,
                           y = Dim.2,
                           text = row.names(spotify_num),
                           color=cluster)) +
                geom_point() +
                theme_void() +  # Remove all axes and labels
                theme(legend.position = "none") +
                scale_colour_brewer(palette="Set2")


ggplotly(cluster_plot, tooltip = "text")

```

